---
title: 'Exploratory Analysis by Aaron'
date: "August 17, 2017"
geometry: margin=1.5cm
output:
  pdf_document: default
  html_document: default
---

##Summary of Findings from EDA

Types of Variables:

1. Rates, averages, and probabilities - crmrte, prbarr, prbconv, prbpris, polpc, density, pctmin80, pctymle
2. $ variables - taxpc, wcon, wtuc, wtrd, wfir, wser, wmfg, wfed, wsta, wloc
3. Indicator variables - west, central, urban. No base category in dataset (e.g. non-west/central, rural)
4. Not sure what this is - mix

Potential Transformations:

1. Percentage minority (pctmin80) data is between 0 - 100 whereas other percentage variables is between 0-1; We may want to transform this variable to keep things consistent from an interpretability perspective.
2. Other than this, so far no obvious variable that needs transformation.

Potential Outliers/Data Issues:

1. Row 51 has prbarr > 1.0 which seems suspicious, given prbarr should between 0 and 1.
2. There seems to be 10 rows where prbconv is > 1.0, which again is suspicious, given that prbconv should be between 0 and 1.
3. Average Sentence in Days (avgsen) looks like it should be Average Sentence in Years.
4. Row 81 looks like an outlier where wser is extremely high. 

##Setup
Reading and understanding the data schema:
```{r}
library(corrplot)
library(car)

setwd("C:\\Users\\aayuen\\Documents\\GitHub\\w203_lab4_kka")
data = read.csv("crime.csv")
nrow(data)
colnames(data)
```

##Univariate Variable Analysis

###1. county - County Identifier

County is essentially a unique identifier (no duplicates).
```{r}
length(unique(data$county))
length(data$county)
```

###2. year - Only 87

The dataset only contains data for 1987.
```{r}
unique(data$year)
```

###3. crmrte - Crimes Committed per Person

```{r fig.height=3}
hist(data$crmrte, breaks=50,
     main="Histogram of Crimes Committed per Person",
     xlab="Crimes committed per person")
```


###4. prbarr - Probability of Arrest

Row 51 has prbarr > 1.0 which seems suspicious, given prbarr should between 0 and 1.
```{r fig.height=3}
summary(data$prbarr)
data[data$prbarr > 1,]
hist(data$prbarr, breaks=50,
     main="Histogram of Probability of Arrest",
     xlab="Probability of arrest")
```

###5. prbconv - Probability of Conviction

There seems to be 10 rows where prbconv is > 1.0, which again is suspicious, given that prbconv should be between 0 and 1.
```{r fig.height=3}
summary(data$prbconv)
data[data$prbconv > 1,]
hist(data$prbconv, breaks=50,
     main="Histogram of Probability of Conviction",
     xlab="Probability of conviction")
```

###6. prbpris - Probability of Prison Sentence

```{r fig.height=3}
summary(data$prbpris)
hist(data$prbconv, breaks=50,
     main="Histogram of Probability of Prison Sentence",
     xlab="Probability of prison sentence")
```

###7. avgsen - Average Sentence in Days

Really? Days? This looks more like average sentence in years...
```{r fig.height=3}
summary(data$avgsen)
hist(data$avgsen, breaks=50,
     main="Histogram of Average Sentence in Days",
     xlab="Average sentence in days")
```


###8. polpc - Police per Capita

Data looks expected.
```{r fig.height=3}
summary(data$polpc)
hist(data$polpc, breaks=50,
     main="Histogram of Police per Capita",
     xlab="Police per capita")
```


###9. density - People per Sq. Mile

Data looks expected.
```{r fig.height=3}
summary(data$density)
hist(data$density, breaks=50,
     main="Histogram of People per Sq. Mile",
     xlab="Probability of people per sq. mile")
```

###10. taxpc - Tax Revenue per Capita

Data looks expected.
```{r fig.height=3}
summary(data$taxpc)
hist(data$taxpc, breaks=50,
     main="Histogram of Tax Revenue per Capita",
     xlab="Tax revenue er capita")
```

###11. west - Indicator of Western N.C.

Data looks expected.
```{r fig.height=3}
summary(factor(data$west))
barplot(table(data$west),
        main="Non-Western N.C. vs Western N.C.")
```

###12. central - Indicator of Central N.C.

Data looks expected.
```{r fig.height=3}
summary(factor(data$central))
barplot(table(data$central),
        main="Non-Central N.C. vs Central N.C.")
```

###13. urban - Indicator of whether in SMSA

What does "SMSA" mean? Otherwise, data looks expected.
```{r fig.height=3}
summary(factor(data$urban))
barplot(table(data$urban),
        main="Non-Urban vs Urban")
```

###14. pctmin80 - Percentage of Minority, 1980

Data looks expected. However, data is between 0 - 100 whereas other percentage variables is between 0-1; We may want to transform this variable to keep things consistent from an interpretability perspective.
```{r fig.height=3}
summary(data$pctmin80)
hist(data$pctmin80, breaks=50,
     main="Histogram of Percentage of Minority",
     xlab="Probability of percentage of minority")
```

###15. wcon - Weekly Wage, Construction

Data looks expected.
```{r fig.height=3}
summary(data$wcon)
hist(data$wcon, breaks=50,
     main="Histogram of Weekly Wage, Construction",
     xlab="Weekly wage")
```

###16. wtuc - Weekly Wage, Trans, Util, Communication

Data looks expected.
```{r fig.height=3}
summary(data$wtuc)
hist(data$wtuc, breaks=50,
     main="Histogram of Weekly Wage, Trans, Util, Communication",
     xlab="Weekly wage")
```

###17. wtrd - Weekly Wage, Wholesale, Retail Trade

Data looks expected.
```{r fig.height=3}
summary(data$wtrd)
hist(data$wtrd, breaks=50,
     main="Histogram of Weekly Wage, Wholesale, Retail Trade",
     xlab="Weekly wage")
```

###18. wfir - Weekly Wage, Finance, Insurance, Real Estate

Data looks expected.
```{r fig.height=3}
summary(data$wfir)
hist(data$wfir, breaks=50,
     main="Histogram of Weekly Wage, Finance, Insurance, Real Estate",
     xlab="Weekly wage")
```

###19. wser - Weekly Wage, Service Industry

There's a very extreme outlier here, need to take a closer look.
```{r fig.height=3}
summary(data$wser)
hist(data$wser, breaks=50,
     main="Histogram of Weekly Wage, Service Industry",
     xlab="Weekly wage")
data[data$wser > 2000, ]
```

###20. wmfg - Weekly Wage, Manufacturing

Data looks expected.
```{r fig.height=3}
summary(data$wmfg)
hist(data$wmfg, breaks=50,
     main="Histogram of Weekly Wage, Manufacturing",
     xlab="Weekly wage")
```

###21. wfed - Weekly Wage, Fed Employees

Data looks expected.
```{r fig.height=3}
summary(data$wfed)
hist(data$wfed, breaks=50,
     main="Histogram of Weekly Wage, Fed Employees",
     xlab="Weekly wage")
```


###22. wsta - Weekly Wage, State Employees

Data looks expected.
```{r fig.height=3}
summary(data$wsta)
hist(data$wsta, breaks=50,
     main="Histogram of Weekly Wage, State Employees",
     xlab="Weekly wage")
```

###23. wloc - Weekly Wage, Local Government Employees

Data looks expected.
```{r fig.height=3}
summary(data$wsta)
hist(data$wsta, breaks=50,
     main="Histogram of Weekly Wage, Local Government Employees",
     xlab="Weekly wage")
```

###24. mix - Offense Mix: Face-to-face/Other

Not really sure what this variable means...
```{r fig.height=3}
summary(data$mix)
hist(data$mix, breaks=50,
     main="Histogram of Offense Mix: Face-to-face/Other",
     xlab="Weekly wage")
```

###25. pctymle - Percent Young Male

Data looks expected.
```{r fig.height=3}
summary(data$pctymle)
hist(data$pctymle, breaks=50,
     main="Histogram of Percent Young Male",
     xlab="Percent of Young Male")
```

##Bi-variate Variable Analysis

Here is the correlation plot for the non-indicator variables.

Looking at the plot, looks like the following variables correlate with crmrte highly and positively:

1. density - Makes sense since the crime rate tends to increase in more populated areas.
2. taxpc - Somewhat makes sense since crime rate could increase in areas where there are more tax revenues collected.
3. wage variables - Somewhat makes sense since as wages go up, there may be higher likelihood for crime. 

Looks like the following variables correlate with crmrte highly and negatively:
1. prbarr - Makes sense since if probabily of arrests go down, then there are more criminals out on the streets.
2. prbconv - Makes sense since if probability of convictions go down, then there are more (potential) criminal out on the streets.

```{r fig.height=6}
corrplot(cor(data[ , (names(data) %in% c("crmrte", "prbarr", "prbconv", "prbpris", "polpc", "density", "pctmin80", "pctymle",
                                          "taxpc", "wcon", "wtuc", "wtrd", "wfir", "wser", "wmfg", "wfed", "wsta", "wloc",
                                          "mix"))]), 
                                           method="circle", type="upper")
```

Let's focus more on the variables that correlate highly with crmrte.

```{r}
scatterplotMatrix(data[ , (names(data) %in% c("crmrte", "prbarr", "prbconv", "density",
                                          "taxpc"))])
```


